---
title: "Inference Analysis: the Tooth Growth case"
output:
  pdf_document: default
  html_notebook: default
author: Leandro GARCIA
---


# OVERVIEW

In this project we will look into the Tooth Growth data set available in R package. This dataset contains the response of 2 groups of guinea pigs in the administration of Vitamin C dosis by 2 different ways. The response is the length of odontoblasts (cells responsible for tooth growth). 60 guinea pigs were tested. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice (coded as *OJ*) or ascorbic acid (a form of vitamin C used as a preservative in food items and coded as *VC*).

# About the Dataset
The Data set contains 60 observations and 3 features.An observation represent a subject for the a specific suplement.
The features are :

- Len : Length of odontoblasts. This feature will be named as 'Length'

- dose : Vitamin C dose. This feature will be named as 'Dosage'

- Supp: The administrated supplement (Vitamin C). This feature will be named as 'Supplement'
    
    - OC: for Orange Juice
    
    - VC: Vitamin C in Ascorbic Acid form
- 

# Basic descriptive statics

``` {r}
str(ToothGrowth)
```

``` {r}
unique(ToothGrowth$dose)
```

We have then, 2 main group (OC, VC), and 3 sub-groups by dosage (0.5, 1.0, 2.0). We will then proceed to group them and get the basic statistics:

```{r}

data_df <- ToothGrowth

library(dplyr)
names(data_df) <- c("Length", "Supplement", "Dosage")

summary_df <- group_by(.data = data_df, Supplement, Dosage) %>% summarise_all(.funs = c("Mean" = mean, "SD" = sd, "Var" = var))

summary_df
```

# CONFIDENCE INTERVALS

``` {r}
auxiliary_df <- group_by(.data = data_df, Supplement, Dosage) %>% tally()
auxiliary_df$df <-auxiliary_df$n - 1 

merged_df <- merge(x = summary_df, y = auxiliary_df, by = c("Supplement", "Dosage"))
merged_df

t <- qt(p = 0.95, df=9, lower.tail = TRUE)

merged_df$CIinf <- merged_df$Mean -  t * merged_df$SD / sqrt(merged_df$n)
merged_df$CIsup <- merged_df$Mean +  t * merged_df$SD / sqrt(merged_df$n)
merged_df$Confidence <- 0.90

select(.data=merged_df, Supplement, Dosage, df, Confidence, CIinf, Mean, CIsup)

```

# HYPOTHESIS TESTING
Our hypothesis is that for doses of 2.0 mg, the supplement type doesn't have a great effect on the tooth growth rate. In other words.
              $H_{0}: \mu_{OJ, 2.0} = \mu_{VC, 2.0}$
              
So, our alternative Hipothesis is :
              $H_{0}: \mu_{OJ, 2.0} \neq \mu_{VC, 2.0}$

## ASSUMPTIONS

* All the numerical variables presented here are IID (Independent and Identical Distribuited)

* The observations may not come from the same population, so we are assuming: 

  - A difference variance for each group 

  - A subject only belong to a group.

* A t-test is the most suitable as we couldn't confirm that the distribution for each group is normal (see annexes)


## t.test() in R
Parameters: 
- alternative = "two.sided"

- paired = FALSE

- var.equal = FALSE

- conf.level = 0.95

```{r}
data_df <- mutate(.data = data_df, "Group" = paste(Supplement, as.character(Dosage), sep="_"))
t.test(x = data_df$Length[data_df$Group == "OJ_2"], y = data_df$Length[data_df$Group == "VC_2"], alternative = "two.sided", paired = FALSE, var.equal = FALSE, conf.level = 0.95)


```

# CONCLUSIONS
Even when our exploratory analysis (see ANNEXES) gives the sensation that because that the group OJ_2 and VC_2 overlaps, their means are the same, in reality they are not and there is a difference between the 2 means.


# ANNEXES

## Basic Exploratory Analysis - Checking the distribution on the 6 groups

``` {r}
library(ggplot2)
data_df <- mutate(.data = data_df, "Group" = paste(Supplement, as.character(Dosage), sep="_"))
ggplot(data = data_df, aes(x = Length, fill = Group)) + geom_density(alpha = 0.2)
```

With the quantity of subjects, it is not clear if they follow a Normal distribution for the 6 groups, therefore we cannot easily kept the idea that the variables are 'normal distributed'.

## Basic Exploratory Analysis - Checking the distribution on the 2 groups

``` {r}
ggplot(data = data_df, aes(x = Length, fill = Supplement)) + geom_density(alpha = 0.2)

```
We can neigher say that the 2 groups are normal distributed.


# REFERENCE
* https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/ToothGrowth.html