---
title: "Central Limit Theorem applied to Exponential Distribution"
output:
  pdf_document: default
  html_notebook: default
author: Leandro GARCIA
---


# OVERVIEW

In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. 

# PROCEDURE
The exponential distribution is simulated in R with `rexp(n, lambda)` where lambda is the rate parameter. The _*mean ($\mu$)*_ of exponential distribution is *$\frac{1}{\lambda}$* and the _*standard deviation ($\sigma$)*_ is also *$\frac{1}{\lambda}$*. For all the simulation, lambda is set to 0.2. We will simulate 1000 sample means, using 40 exponentials. 

# THEORETICAL VALUES
In this section we calculate the theoretical values for the exponential distribution distribution


## Theoretical Mean: ( $\mu_{t} = \frac{1}{\lambda}$ ) 

```{r}
lambda_t <- 0.2 # Value for simulation
mu_t <- 1/lambda_t
mu_t # 5.0
```
Theoretical mean ( $\mu_{t} = 5.0$ ) 


## Theoretical Standard Desviation ( $\sigma_{t} = \frac{1}{\lambda}$ ) 

```{r}
# lambda_t <- 0.2
sigma_t <- 1/lambda_t
sigma_t # 5.0
```
Theoretical Standard Deviation: ( $\sigma_{t} = 5.0$ ) 


## Theoretical Variance ( $Var_{t} = \sigma^2$ ) 

```{r}
# sigma_t <- 5.0
var_t <- sigma_t^2 
var_t # 25.0
```
Theoretical Variance: ( $Var_{t} = 25.0$ ) 

## Theoretical Standard Error of the mean : ( $SE_{\overline{X}} = \frac{\sigma}{\sqrt(n)}$ ) 

```{r}
sample_size <- 40
SE_t <- sigma_t / sqrt(sample_size) 
SE_t # 25.0
```
Theoretical Standar Error of the mean : ( $SE_{\overline{X}} = 0.79$ )


# PRACTICAL VALUES
In this section we calculate the random values of the distribution, and then we will proceed to retrieve the initial parameter (inference).

## SIMULATING 

Getting the values to Simulate 1000 times 40 $X_{1 \dots 40}$, and then obtain 1000 $\overline{X}_{i}$ or in other words $\overline{X}_{1  \dots 1000}$

```{r}
# Parameters for simulation
set.seed(1234)
means_size <- 1000
sample_size <- 40

# Generate the samples
data_list = rexp(means_size * sample_size, rate = lambda_t) # to be used for plotting
data_matrix <- matrix( data = data_list , nrow = means_size, ncol = sample_size)

# Getting the means for all the simulation
means <- apply(X = data_matrix, MARGIN = 1, FUN = mean )

# Simulated means based on the parameters
str(means)

```

## Practical mean: ( $\mu_{p} = \sum_{i=1}^{1000}\frac{\overline{X}_{i}}{n}$ )

```{r}
?sd
mu_p <- mean(means)
mu_p


```

Practical mean: ( $\mu_{p} = 4.97$ )

## Practical Standard Error of the mean: ( $SE_{\overline{X}} = \sigma_{\overline{X}}$ ) 
```{r}
?sd
SE_p <- sd(means)
SE_p
```

Practical Standard Error of the mean: ( $SE_{\overline{X}} = 0.77$ )

## Practical Population Standard Deviation: ($\sigma_{p} = SE_{\overline{X}} * \sqrt(n)$ ) 
```{r}
?sd
sigma_p <- SE_p * sqrt(sample_size)
sigma_p
```

Practical Population Standard Deviation: ($\sigma_{p} = 4.88$ )

## Practical Population Variance: ( $Var_{p} = \sigma^2$ ) 
```{r}
?sd
var_p <- sigma_p ^ 2
var_p
```

Practical Population Variance: ( $Var_{p} = 23.8$ )


# THEORETICAL VALUES Vs. PRACTICAL VALUES

```{r}
library(ggplot2)
means_df<- data.frame(value = means, type = rep("Mean distribution", length(means)))
data_df <- data.frame(value = data_list, type = rep("data distributiont", length(data_list)))
all_data <- rbind(means_df, data_df)

ggplot(data = all_data, aes(x = value, fill = type)) + geom_density(alpha = 0.2)

```

STATISTIC          | THEORETICAL | PRACTICAL
-------------------|-------------|------------
$\mu$              |  5.0        |  4.97
$Var$              | 25.0        |  4.88
$\sigma$           |  5.0        | 23.8
$SE_{\overline{X}}$|  0.79       |  0.77


# CONCLUSIONS:

Using random values of a exponential distribution, we were able to :

- Demonstrate the relationship that exist between the distribution of the average of the means and the population means. 

- Prove that the distribution of the means is a *normal distribution* even when the population distribution is not.

- Confirm that the relationship between the population distribution and means distribution is dictated by the *Central Limit Theorem*. 

```{undefined eval=FALSE, include=FALSE}
# REFERENCES
General references used to build this study:

* https://rpruim.github.io/s341/S19/from-class/MathinRmd.html

* https://www.youtube.com/watch?v=aIPvgiXyBMI

* https://stackoverflow.com/questions/3541713/how-to-plot-two-histograms-together-in-r
```